\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-05-31 jss/cubble-jss.Rmd
also knit the jss version
match jss abstract with the arxiv version

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    # location: Melbourne, Australia
    email: dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College \AND'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolaslangrene@uic.edu.cn
  - name: Patricia Menéndez
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  Spatio-temporal data refer to measurements taken across space and time. In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution. Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects. To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations. This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object. The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data. It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation. The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
keywords:
  # at least one keyword must be supplied
  formatted: [spatial, temporal, spatio temporal, "\\proglang{R}", environmental data, exploratory data analysis]
  plain:     [spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{xcolor}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
bibliography: ../references.bib
---

>>>>

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    # location: Melbourne, Australia
    email: dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College \AND'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolaslangrene@uic.edu.cn
  - name: Patricia Menéndez
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
    Multivariate spatio-temporal data refers to multiple measurements taken across space and time. For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time. However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations. In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components.  The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis. In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling. The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures. The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
keywords:
  # at least one keyword must be supplied
  formatted: [spatial, temporal, spatio temporal, "\\proglang{R}", environmental data, exploratory data analysis]
  plain:     [spatial, temporal, spatio temporal, R, environmental data, exploratory data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{xcolor}
  \newcommand{\sherry}[1]{\textcolor{magenta}{#1}}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
bibliography: ../references.bib
---

### 2022-05-31 jss/cubble-jss.tex
match jss abstract with the arxiv version

\Abstract{
Spatio-temporal data refer to measurements taken across space and time. In practice, spatio-temporal data can be decomposed into a spatial and temporal component: at one time, we would select a spatial location and inspect the temporal trend; at other time, we might select one or multiple time value(s) and explore the spatial distribution. Ideally, we could make multiple maps and multiple time series to explore these together, however, doing all of these actions is complicated when data arrive fragmented in multiple objects. To make it easy to do all these tasks, ideally spatial and temporal variables are in a single data object that we can slice and dice in different ways to conduct different visualisations. This work suggests a new data structure, \code{cubble}, that uses a nested form and a long form to organise spatial and temporal variables in a single data object. The new data structure ensures that data in the two forms are synchronised while been flexible to explore the two aspect of spatio-temporal data. It also provides capability for handling data with hierarchical structure, matching data from multiple sources, constructing interactive graphics, and performing spatio-temporal transformation. The paper will demonstrate \pkg{cubble} with examples using Australian climate weather stations, river level data, and climate reanalysis (ERA5) data.
}

>>>>

\Abstract{
Multivariate spatio-temporal data refers to multiple measurements taken across space and time. For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time. However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations. In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components. The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis. In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling. The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures. The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
}
